---
title: "Data Transformation - Part IV"
output:
  html_document:
    df_print: paged
Name: THI NGOC DIEM NGUYEN
ID: 0778061
Week: 7
---

```{r set up, include=FALSE}
#knitr::opts_chunk$set(eval = FALSE)
```

## Load packages

```{r warning = FALSE, message = FALSE}
library(tidyverse)
library(gapminder)
library(nycflights13)
```

## The ''summarize' verb

'summarize' takes a dataset with $n$ observations, computes the requested summaries, and returns a dataset with 1 observation. Here we calculate the average life expectancy.

```{r}
summarize(gapminder, avg_life = mean(lifeExp))
```

1. Find the average GDP per capita.

```{r}
summarize(gapminder, avg_GDP = mean(gdpPercap))
```

2. Find the average population.

```{r}
summarize(gapminder, avg_pop = mean(pop))
```

Some other functions that take a vector of values and return a single value that work well with 'summarize' are: first, last, min, max, nth, n, n_distinct, median, var, sd, IQR.

3. Use 'n_distinct' to find the number of continents in the data frame.

```{r}
summarize(gapminder, n_distinct(continent))
```

4. Use 'last' to find the last country listed in the data frame.

```{r}
summarize(gapminder, last(country))
```

We can also use 'summarize_all' to act on multiple columns at once:

```{r}
gapminder %>% 
  select(lifeExp, pop, gdpPercap) %>% 
  summarize_all(funs(mean, median))
```

## Pipe Operator

From here on we will start using the pipe operator.

## Grouping

Things get more interesting when we start using 'group_by' to form groups within our data frame and then perform analysis group by group.

5. In the console, type 'gapminder' and then type 'gapminder %>% group_by(continent)'. Note the information about the grouping by continent at the top of the data frame. (This metadata does not get displayed in a notebook.)

Here we count the number of observations per continent:

```{r}
gapminder %>% group_by(continent) %>% 
  summarize(n = n())
```

And here we find the number of distinct countries per continent:

```{r}
gapminder %>% group_by(continent) %>% 
  summarize(n_countries = n_distinct(country))
```

6. Find the average life expectancy per year by continent.

```{r}
gapminder %>% 
  group_by(continent, year) %>% 
  summarize(avg_life_exp = mean(lifeExp))
```

7. Find the highest GDP per capita by year and arrange in descending order.

```{r}
gapminder %>% 
  group_by(year) %>% 
  summarize(max_GDP = max(gdpPercap)) %>% 
  arrange(desc(max_GDP))
```

8. Find the lowest GDP per capita by year and arrange in ascending order.

```{r}
gapminder %>% 
  group_by(year) %>% 
  summarize(min_GDP = min(gdpPercap)) %>% 
  arrange(min_GDP)
```

## Let’s switch over to 'nycflights13'.

9. Which airline had the highest average departure delay?

```{r}
flights %>% 
  group_by(carrier) %>% 
  summarize(avg_delay = mean(dep_delay, na.rm = TRUE)) %>% 
  arrange(desc(avg_delay))
```

10. Which month had the most flights?

```{r}
flights %>% 
  group_by(month) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n))
```

11. Which month had the most on-time (or early) departures?

```{r}
flights %>% filter(dep_delay <= 0) %>%
  group_by(month) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n))
```

12. Which of the 3 New York City airports had the highest average departure delay?

```{r}
flights %>% filter(!is.na(dep_delay)) %>%
  group_by(origin) %>% 
  summarize(avg_dep_delay = mean(dep_delay)) %>% 
  arrange(desc(avg_dep_delay))
```

13. For the airport identified above, was there a particular month that had a high average departure delay? Or were all the months about the same?

```{r}
flights %>% filter(origin == 'EWR', !is.na(dep_delay)) %>%
  group_by(month) %>% 
  summarize(avg_dep_delay = mean(dep_delay)) %>% 
  arrange(desc(avg_dep_delay))
```

14. Which destination airport had the highest arrival delay? Any particular months stand out as the cause? Any particular airline?

```{r}
flights %>% filter(!is.na(arr_delay)) %>%
  group_by(dest) %>% 
  summarize(avg_arr_delay = mean(arr_delay)) %>% 
  arrange(desc(avg_arr_delay))
```

```{r}
flights %>% filter(!is.na(arr_delay)) %>%
  group_by(dest, month) %>% 
  summarize(avg_arr_delay = mean(arr_delay)) %>% 
  arrange(dest, desc(avg_arr_delay))
```

```{r}
flights %>% filter(!is.na(arr_delay)) %>%
  group_by(dest, carrier) %>% 
  summarize(avg_arr_delay = mean(arr_delay)) %>% 
  arrange(dest, desc(avg_arr_delay))
```

15. Find the first and last departure for each day.

```{r}
flights %>% filter(!is.na(dep_delay) & !is.na(arr_delay)) %>%
  group_by(year, month, day) %>% 
  summarize(first = min(dep_time), last = max(dep_time))
```

Sometimes you will want to remove the grouping. To do that we can do the following:

```{r}
flights %>% 
  group_by(month) %>% 
  summarize(n = n()) %>% 
  ungroup()
```

## Grouped mutates

Sometimes you don’t want to collapse the n rows for each group into one row. You want to keep your groups, but compute within them.

Find the top two carriers by origin airport ranked according to average departure delay:

```{r}
flights %>% group_by(origin, carrier) %>% 
  summarize(avg_dep_delay = mean(dep_delay, na.rm = TRUE)) %>% 
  mutate(rank = rank(desc(avg_dep_delay))) %>% 
  filter(rank < 3) %>% 
  arrange(origin, rank)
```

16. For each destination, compute the total minutes of delay. For each flight, compute the proportion of the total delay for its destination.

```{r}
flights %>% filter(!is.na(arr_delay) & arr_delay > 0) %>% 
  group_by(dest) %>% 
  mutate(arr_delay_total = sum(arr_delay), arr_delay_prop = arr_delay / arr_delay_total) %>% 
  select(dest, arr_delay_total, arr_delay_prop) %>% 
  arrange(dest)
```

## Counting

Here we count the number of flights per month.

```{r}
flights %>% group_by(month) %>% 
  summarize(n = n())
```

17. Count the number of flights per carrier.

```{r}
flights %>% group_by(carrier) %>% 
  summarize(n = n())
```

18. Count the number of flights per origin.

```{r}
flights %>% group_by(origin) %>% 
  summarize(n = n())
```

19. Count the number of flights by carrier, destination, and month.

```{r}
flights %>% group_by(carrier, dest, month) %>% 
  summarize(n = n())
```

Each time we apply summarize we remove one layer of grouping.

```{r}
flights %>% group_by(carrier, dest, month) %>% 
  summarize(n = n()) %>% 
  summarize(n = n()) %>% 
  summarize(n = n())
```

Counts are so useful that 'dplyr' has a helper function that does the grouping and counting. Here we 'count' the number of flights per plane using 'count' instead of grouping by 'tailnum' and then using 'summarize(n = n())'.

```{r}
flights %>% 
  filter(!is.na(dep_delay) & !is.na(arr_delay)) %>% 
  count(tailnum)
```

## Why do we care about counting

20. Exclude cancelled flights (i.e., filter out observations where
'dep_delay' and 'arr_delay' is missing. Save the results in a variable called 'not_cancelled'.

```{r}
not_cancelled <- flights %>% 
  filter(!is.na(dep_delay) & !is.na(arr_delay))
```

21. Determine the average arrival delay and counts by tailnum for these flights and store in a variable called 'delays'.

```{r}
delays <- not_cancelled %>% group_by(tailnum) %>%
  summarise(avg_arr_delay = mean(arr_delay), n = n())
delays
```

22. Create a scatter plot of counts vs average arrival delays. Use 'shape = 21' and 'alpha = 0.25'.

```{r}
ggplot(delays, aes(x = n, y = avg_arr_delay)) + geom_point(shape = 21, alpha = 0.25)
```

23. How does average arrival delay change as the number of flights a plane takes increases.

More counts, less delay
Less counts, more delay
-> Average arrival delay decreases when the number of flights a plane takes increases.

24. Re-create the scatter plot above but only consider planes that have flown at least 25 flights.

```{r}
not_cancelled %>% group_by(tailnum) %>%
  summarise(avg_arr_delay = mean(arr_delay), n = n()) %>%
  filter(n > 25) %>%
  ggplot(aes(x = n, y = avg_arr_delay)) + geom_point(shape = 21, alpha = 0.25)
```

If we are drawing conclusions from the data we need to make sure we are basing them on a sufficient amount of data.

