---
title: "Data Transformation - Part III"
output:
  html_document:
    df_print: paged
Name: THI NGOC DIEM NGUYEN
ID: 0778061
Week: 7
---

```{r set up, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

## Load packages

```{r warning = FALSE, message = FALSE}
library(tidyverse)
library(gapminder)
library(nycflights13)
```

## The 'mutate' verb

'mutate' allows us to create new variables or modify existing one.

1. Create a new variable called 'GDP', which is the product of 'pop' and 'gdpPercap'. Check the data frame to make sure your new variable was created.

```{r}
GDP <- mutate(gapminder, GDP = pop * gdpPercap)
GDP
```

2. Create another new variable called 'GDP_norm', which is the 'GDP' in billions of dollars. Check the data frame to make sure your new variable was created.

```{r}
GDP_norm <- mutate(GDP, GDP_norm = GDP / 1000000000 )
GDP_norm
```

3. Modify the 'pop' variable so it expressed units of millions of people. Check the data frame to make sure that the 'pop' variable was modified.

```{r}
mutate(gapminder, pop = pop / 1000000 )
```

## The 'transmute' verb

This verb allows you to create new variables but removes the existing ones.

4. Use 'transmute' to create a 'population' variable measured in units of millions of people. Check that the 'population' variable was created and that the other variables were removed.

```{r}
transmute(gapminder, population = pop)
```

## Rename

This verb allows you to rename variables.

Compare the columns in

```{r}
mutate(gapminder, population = pop)
```

to the columns in

```{r}
rename(gapminder, population = pop)
gapminder
```

## The 'arrange' verb

This verb allows you to reorder the rows of your data frame.

5. Reorder the rows in increasing order based on 'lifeExp'.

```{r}
arrange(gapminder, lifeExp)
```

6. See if you can reorder based on 'gdpPercap' but in decreasing order.

```{r}
arrange(gapminder, desc(gdpPercap))
```

7. Can you use this function to reorder based on 'year' and 'pop'? What happens if you reverse the order of the variables?

```{r}
arrange(gapminder, year, pop)
```

```{r}
arrange(gapminder, pop, year)
```
-> The order depends on what column we are putting first in the chunk.

## Using the 'nycflights13' data:

# Install a New Package

```{r eval=FALSE}
install.packages('nycflights13')
```

# Load Packages

```{r warning = FALSE, message = FALSE}
library(nycflights13)
flights
```

8. Sort flights to find the most delayed flights.

```{r}
arrange(flights, desc(dep_delay))
```

9. Find the flights that left earliest.

```{r}
arrange(flights, dep_delay)
```

10. Sort flights to find the fastest flights.

```{r}
arrange(flights, air_time)
```

11. Which travelled the shortest?

```{r}
arrange(flights, distance)
```

Which flights travelled the longest?

```{r}
arrange(flights, desc(distance))
```

## Section 5.5.2 Exercises

```{r}
name_of_function <- function(x) {}
```

12. Currently 'dep_time' and 'sched_dep_time' are convenient to look at, but hard to compute with because they’re not really continuous numbers. Convert them to a more convenient representation of number of minutes since midnight.

consider the first value in 'dep_time': 517. This actually means 5:17, or 17 minutes after 5pm. if we do some modular arithmetic we can separate the hours from the minutes. We can then use these values to convert 'dep_time' to minutes from midnight.

# Extract the hours:

```{r}
517 %/% 100
```

# Extract the minutes:

```{r}
517 %% 100
```

If we convert the hours to minutes and add both together we should get total minutes from midnight:

```{r}
(517 %/% 100) * 60 + (517 %% 100)
```

See how this works on our minimum and maximum values of 'dep_time':

```{r warning = FALSE, message = FALSE}
library(nycflights13)
flights
```

```{r}
summary(flights$dep_time)
```

Our minimum is 1, which represents 12:01 (one minute after midnight), so when we convert to minutes from midnight we should still have a value of 1:

```{r}
(1 %/% 100) * 60 + (1 %% 100)
```

Our maximum is 2400, which represents 12:00 (midnight), so when we convert to minutes from midnight we should get a value of 0:

```{r}
(2400 %/% 100) * 60 + (2400 %% 100)
```

We do not get 0. Since this is maximum value, we can do one more modular division to fix this issue:

```{r}
((2400 %/% 100) * 60 + (2400 %% 100)) %% 1440
```

```{r eval=FALSE}
install.packages('nycflights13')
```

```{r warning = FALSE, message = FALSE}
library(tidyverse)
library(nycflights13)
flights
```

We could use the above formula directly in a mutate statement.

```{r}
mutate(flights, dep_time_in_mins = ( (dep_time %/% 100) * 60 + (dep_time %% 100) ) %% 1440)
```


A better way is to use function:

```{r}
time_in_mins <- function(x) {((x %/% 100) * 60 + (x %% 100)) %% 1440}
```

Test new function with a few values:

```{r}
print(time_in_mins(1))
print(time_in_mins(2400))
print(time_in_mins(752))
```

Now we can re-write what we had above:

```{r}
mutate(flights, dep_time_in_mins = time_in_mins(dep_time))
```

13. Compare 'air_time' with 'arr_time' - 'dep_time'. What do you expect to see? What do you see? What do you need to do to fix it?

It would be logical to assume that 'arr_time' - 'dep_time' = 'air_time'.

Let's check:

First, 'arr_time' and 'dep_time' need to be converted as we did in the previous question. Then we need to calculate the difference. Then we select only the columns of interest to make viewing simpler.

calculate the difference between 'air_time' and our computed_air_time = 'arr_time' - 'dep_time'

Plot a histogram of the differences (which we think should be 0)

```{r}
flights %>% mutate(arr_time = time_in_mins(arr_time), 
                   dep_time = time_in_mins(dep_time), 
                   computed_air_time = arr_time - dep_time,
                   difference = air_time - computed_air_time) %>% 
  ggplot(aes(x = difference)) + geom_histogram(colour = 'black', fill = 'lavender', binwidth = 50)
```

Possible issues could be:
- time zones
- arrival and departure are on different days
- air_time does not include taxi and runway time

14. Compare 'dep_time', 'sched_dep_time', and 'dep_delay'. How would you expect those three numbers to be related?

Similar to above but assuming: 'dep_time' = 'sched_dep_time' + 'dep_delay'

```{r}
flights %>% mutate(dep_time = time_in_mins(dep_time), 
                   sched_dep_time = time_in_mins(sched_dep_time),
                   diff_sched_dep = (dep_time - dep_delay) - sched_dep_time) %>% 
  ggplot(aes(x = diff_sched_dep)) + geom_histogram(colour = 'black', fill = 'lavender', binwidth = 50)
```

Possible issues could be:
- There are some flights scheduled near midnight that get delayed and leave next day.






