---
title: "Exploratory Data Analysis - Part I"
output:
  html_document:
    df_print: paged
Name: THI NGOC DIEM NGUYEN
ID: 0778061
Week: 8
---

```{r set up, include=FALSE}
#knitr::opts_chunk$set(eval = FALSE)
```

## Install a New Package

```{r eval=FALSE}
install.packages('here')
```

## Load necessary packages

```{r warning = FALSE, message = FALSE}
library(tidyverse)
library(here)
```

## Load the data

Data set is the Bwages.csv file (Wages in Belgium), as documented at http://vincentarelbundock.github.io/Rdatasets/datasets.html

1. Run the code below to read the data set in from the web.

```{r}
wages <- read_csv("http://vincentarelbundock.github.io/Rdatasets/csv/Ecdat/Bwages.csv")
```

2. Take a moment to look at the documentation describing the data set: http://vincentarelbundock.github.io/Rdatasets/doc/Ecdat/Bwages.html

## Initial exploration

3. Take an initial look at the data set. How many observations? How many variables?

```{r}
glimpse(wages)
wages
```

-> There are 1,472 observations and 5 variables.

4. What variables do we have? What data types are they stored as? What data types should they be? Document any changes that need to be made.

- X1 - integer - doesn't seem to be a variable
- wage - double - continuous numeric
- educ - integer - categorical (factor)
- exper - integer - continuous numeric
- sex - character - categorical (factor)

5. Is there any missing data? In which variables? What percent of the variable does the missing data represent? How should we fix this problem?

```{r}
summary(wages)
```
```{r}
sum(is.na(wages$sex))
```

- Is there any missing data? -> Yes
- In which variables? -> The 'sex' variable
- What percent of the variable does the missing data represent? -> 100%
- How should we fix this problem? -> The 'sex' variable has no data, we should remove it.

## Data clean up

6. Based on your discoveries above, clean up the data for analysis.

- remove the 'X1' variable as it's just a row number
- remove the 'sex' variable as it contains no data
- convert the 'educ' variable to a factor

```{r}
wages_clean <- wages %>% select(-X1, -sex) %>% mutate(educ = factor(educ))
wages_clean
```

## Univariate analysis

In this section you will need to consider one variable at a time.

### The 'wage' variable

7. Give a brief description of this variable.

'wage': gross hourly wage rate in euro

8. From a statistics point of view, what data type is the 'wage' variable.

continuous numeric

9. Create a histogram showing the distribution of the 'wage' variable. Use both 'colour' and 'fill' attributes so you can clearly see the individual bins in the plot.

```{r}
ggplot(wages_clean, aes(x = wage)) + geom_histogram(fill = 'skyblue', colour = 'black')
```

10. Calculate the maximum and minimum values of the 'wage' variable.

```{r}
print(min(wages_clean$wage))
print(max(wages_clean$wage))
```

11. Use the maximum and minimum values of 'wage' to determine the 'range' of this variable.

```{r}
wage_range <- max(wages_clean$wage) - min(wages_clean$wage)
wage_range
```

12. Divide the range by 30 to get a value for the default 'binwidth' used by 'geom_histogram'.

```{r}
default_bin <- wage_range / 30
default_bin
```

13. Create new histograms by adjusting the default 'binwidth': try 2 values below the default and 2 values above.

```{r}
ggplot(wages_clean, aes(x = wage)) + geom_histogram(fill = 'skyblue', colour = 'black', binwidth = 2)
```

```{r}
ggplot(wages_clean, aes(x = wage)) + geom_histogram(fill = 'skyblue', colour = 'black', binwidth = 3)
```

14. Based on your 'binwidth' exploration above, choose the 'binwidth' that gives the best view of the shape of the distribution of values for 'wage'.

```{r}
ggplot(wages_clean, aes(x = wage)) + geom_histogram(fill = 'skyblue', colour = 'black', binwidth = 2)
```

15. Describe the shape and skewness of this distribution.

Unimodal and right skewed

16. Calculate the average wage and plot this as a vertical red line on your histogram.

```{r}
avg_wage <- mean(wages_clean$wage)
avg_wage
ggplot(wages_clean, aes(x = wage)) + 
  geom_histogram(fill = 'skyblue', colour = 'black', binwidth = 2) + 
  geom_vline(xintercept = avg_wage, colour = 'red', size = 1.5)

```

17. Calculate the median wage and plot this as a vertical blue line on your histogram.

```{r}
med_wage <- median(wages_clean$wage)
med_wage
ggplot(wages_clean, aes(x = wage)) + 
  geom_histogram(fill = 'skyblue', colour = 'black', binwidth = 2) + 
  geom_vline(xintercept = avg_wage, colour = 'red', size = 1.5) +
  geom_vline(xintercept = med_wage, colour = 'blue', size = 1.5)
```

18. Is the mean higher, about the same as, or lower than the median? Use the skewness of the distribution to explain your answer.

The mean is higher than the median which is what we would expect for a right skewed distribution.

19. Are there any possible 'outliers' in the data? If yes, use 'dplyr' verbs to take a look at these possible outlier values. Given what you know, are they reasonable or are they obviously errors? Explain.

```{r}
wages_clean %>% filter(wage > 35)
```

A wage of 50 euros an hour is about 'r format(50*1.5*40*50, scientific = FALSE)' Canadian dollars a year. This is well above average but nothing crazy seems like the outliers could be reasonable data points.

20. Would it be better to use the 'mean' or 'median' as a measure of central tendency for the 'wage' variable? Why?

Median since the distribution is right skewed.

21. What would be a robust 'measure of spread' for this variable? Calculate its value.

interquartile range (IQR)

```{r}
IQR(wages_clean$wage)
```

### The 'exper' variable

22. Repeat the steps carried out for the 'wage' variable but this time use the 'exper' variable.

22.1. Give a brief description of this variable.

'exper': years of experience

22.2. From a statistics point of view, what data type is the 'exper' variable.

continuous numeric

22.3. Create a histogram showing the distribution of the 'exper' variable. Use both 'colour' and 'fill' attributes so you can clearly see the individual bins in the plot.

```{r}
ggplot(wages_clean, aes(x = exper)) + geom_histogram(fill = 'skyblue', colour = 'black')
```

22.4. Calculate the maximum and minimum values of the 'exper' variable.

```{r}
print(min(wages_clean$exper))
print(max(wages_clean$exper))
```

22.5. Use the maximum and minimum values of 'exper' to determine the 'range' of this variable.

```{r}
exper_range <- max(wages_clean$exper) - min(wages_clean$exper)
exper_range
```

22.6. Divide the range by 30 to get a value for the default 'binwidth' used by 'geom_histogram'.

```{r}
default_bin <- exper_range / 30
default_bin
```

22.7. Create new histograms by adjusting the default 'binwidth': try 2 values below the default and 2 values above.

```{r}
ggplot(wages_clean, aes(x = exper)) + geom_histogram(fill = 'skyblue', colour = 'black', binwidth = 2)
```

```{r}
ggplot(wages_clean, aes(x = exper)) + geom_histogram(fill = 'skyblue', colour = 'black', binwidth = 3)
```

22.8. Based on your 'binwidth' exploration above, choose the 'binwidth' that gives the best view of the shape of the distribution of values for 'exper'.

```{r}
ggplot(wages_clean, aes(x = exper)) + geom_histogram(fill = 'skyblue', colour = 'black', binwidth = 2)
```

22.9. Describe the shape and skewness of this distribution.

Multimodal and right skewed

22.10. Calculate the average 'exper' and plot this as a vertical red line on your histogram.

```{r}
avg_exper <- mean(wages_clean$exper)
avg_exper
ggplot(wages_clean, aes(x = exper)) + 
  geom_histogram(fill = 'skyblue', colour = 'black', binwidth = 2) + 
  geom_vline(xintercept = avg_exper, colour = 'red', size = 1.5)

```

22.11. Calculate the median 'exper' and plot this as a vertical blue line on your histogram.

```{r}
med_exper <- median(wages_clean$exper)
med_exper
ggplot(wages_clean, aes(x = exper)) + 
  geom_histogram(fill = 'skyblue', colour = 'black', binwidth = 2) + 
  geom_vline(xintercept = avg_exper, colour = 'red', size = 1.5) +
  geom_vline(xintercept = med_exper, colour = 'blue', size = 1.5)
```

22.12. Is the mean higher, about the same as, or lower than the median? Use the skewness of the distribution to explain your answer.

The mean is higher than the median which is what we would expect for a right skewed distribution.

22.13. Are there any possible 'outliers' in the data? If yes, use 'dplyr' verbs to take a look at these possible outlier values. Given what you know, are they reasonable or are they obviously errors? Explain.

```{r}
wages_clean %>% filter(exper > 45)
```

A person with exper of 50 year is beyond working age. 

22.14. Would it be better to use the 'mean' or 'median' as a measure of central tendency for the 'exper' variable? Why?

Median since the distribution is right skewed.

22.15. What would be a robust 'measure of spread' for this variable? Calculate its value.

interquartile range (IQR)

```{r}
IQR(wages_clean$exper)
```

### The 'educ' variable

23. Give a brief description of this variable.

'educ': education level from 1 [low] to 5 [high]

24. From a statistics point of view, what data type is the 'educ' variable.

categorical (factor)

25. What are the unique values that the variable 'educ' can take on?

```{r}
unique(wages_clean$educ)
```

26. Create a bar plot of the counts for the various levels of 'educ'.

```{r}
ggplot(wages_clean, aes(x = educ)) + geom_bar(fill = 'skyblue', colour = 'black')
```

27. Using 'dplyr', manually calculate these counts and verify that the results are the same as what is shown in the previous bar plot.

```{r}
wages_clean %>% count(educ)
```

28. Create a bar plot of the proportions for the various levels of 'educ'.

```{r}
ggplot(wages_clean, aes(x = educ, y = ..prop.., group = 1), stat = 'count') + geom_bar(fill = 'skyblue', colour = 'black')
```

29. Using 'dplyr', manually calculate these proportions and verify that the results are the same as what is shown in the previous bar plot.

```{r}
wages_clean %>% group_by(educ) %>% 
  summarise(n = n()) %>%
  mutate(prop = n / sum(n))
```

30. Sum up the proportions calculated above. What value do you get? Does that make sense?

```{r}
wages_clean %>% group_by(educ) %>% 
  summarise(n = n()) %>%
  mutate(prop = n / sum(n)) %>%
  summarise(total = sum(prop))
```

This value means we have done it correctly. 1 is total of the proportions, it totally makes sense.

31. Which level of 'educ' has the most counts? the fewest?

```{r}
wages_clean %>% group_by(educ) %>% 
  summarise(n = n()) %>%
  mutate(prop = n / sum(n)) %>%
  arrange(desc(n))
```

-> level of 'educ' 3 has the most counts.

```{r}
wages_clean %>% group_by(educ) %>% 
  summarise(n = n()) %>%
  mutate(prop = n / sum(n)) %>%
  arrange(n)
```

-> level of 'educ' 1 has the fewest counts.



