---
title: "Exploratory Data Analysis - Part II"
output:
  html_document:
    df_print: paged
  pdf_document: default
Name: THI NGOC DIEM NGUYEN
ID: 0778061
Week: 8
---

```{r set up, include=FALSE}
#knitr::opts_chunk$set(eval = FALSE)
```

## Install a New Package: gridExtra

1. In the console (not in an R notebook or R script file), run install.packages('gridExtra') to install a package that will let us combine plots.

```{r eval=FALSE}
install.packages('gridExtra')
```

## Load packages

```{r warning = FALSE, message = FALSE}
library(tidyverse)
library(here)
library(gridExtra)
```

## Univariate analysis (cont’d)

### Measure of central tendency

For numeric variables, the measure of central tendency will give us an idea of where the center of the distribution is; that is, the value around which our data our data values are centered. For symmetric distributions, we can use the mean:

$$
\bar{x} = \frac{1}{n}\sum_{i = 1}^n x_i
$$

For skewed data, we can use the median, or middle value of the data, if the number of data points is even, or the average of the middle two numbers if the number of data points is odd:

```{r}
print(median(c(1, 2, 3, 4, 5)))
print(median(c(1, 2, 3, 4, 5, 6)))
```

For repeated values, it works as follows:

```{r}
median(c(1, 1, 2, 2, 2, 3, 4))
```

### Measure of spread

Variation describes the behavior of values for a single variable.

The variance is the mean squared deviation and is calculated using the formula

$$
s^2 = \sum_{i = 1}^n \frac{(x_i - \bar{x})^2}{n -1} = \frac{(x_1 - \bar{x})^2 + (x_2 - \bar{x})^2 + ... + (x_n - \bar{x})^2}{n - 1}
$$

And the standard deviation is

$$
s = \sqrt{s^2}
$$

If the data is symmetric, the standard deviation gives us a measure of the spread of the data about the center.

If the data is skewed, then the interquartile range (IQR) gives us a measure of spread of the data about the center. Remember that the IQR is the value of the third quartile (or 75th percentile) minus the first quartile (or 25th percentile) and thus tells us how spread out the middle 50% of the data is.

```{r}
IQR(c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10))
```

### Exploring the mean and standard deviation

In this section we compare a variety of means and standards deviations to get a better idea of what these numbers tell us.

NOTE: the 'gridExtra' package needs to be installed.

#### Same mean but different standard deviation

```{r}
norms <- tibble(x_wider = rnorm(1000, mean = 0, sd = 10), 
                x_narrower = rnorm(1000, mean = 0, sd = 1))


p1 <- ggplot(norms, aes(x = x_wider)) +
  geom_histogram(fill = 'skyblue2', colour = 'black', binwidth = 5) +
  coord_cartesian(xlim = c(-30, 30)) +
  ggtitle("Mean = 0 and Standard deviation = 10")


p2 <- ggplot(norms, aes(x = x_narrower)) +
  geom_histogram(fill = 'skyblue2', colour = 'black', binwidth = 0.5) +
  coord_cartesian(xlim = c(-30, 30)) +
  ggtitle("Mean = 0 and Standard deviation = 1")


grid.arrange(p1, p2)
```

#### Different mean but same standard deviation

```{r}
norms_2 <- tibble(x_wider = rnorm(1000, mean = 0, sd = 10), 
                  x_wider_2 = rnorm(1000, mean = 15, sd = 10))


p3 <- ggplot(norms_2, aes(x = x_wider)) +
  geom_histogram(fill = 'skyblue2', colour = 'black', binwidth = 5) +
  coord_cartesian(xlim = c(-30, 45)) +
  ggtitle("Mean = 0 and Standard deviation = 10")


p4 <- ggplot(norms_2, aes(x = x_wider_2)) +
  geom_histogram(fill = 'skyblue2', colour = 'black', binwidth = 5) +
  coord_cartesian(xlim = c(-30, 45)) +
  ggtitle("Mean = 15 and Standard deviation = 10")


grid.arrange(p3, p4)
```

#### Different mean and different standard deviation

```{r}
norms_3 <- tibble(x_wider = rnorm(1000, mean = 15, sd = 10), 
                  x_narrower = rnorm(1000, mean = 28, sd = 1))


p5 <- ggplot(norms_3, aes(x = x_wider)) +
  geom_histogram(fill = 'skyblue2', colour = 'black', binwidth = 5) +
  coord_cartesian(xlim = c(-20, 45)) +
  ggtitle("Mean = 15 and Standard deviation = 10")


p6 <- ggplot(norms_3, aes(x = x_narrower)) +
  geom_histogram(fill = 'skyblue2', colour = 'black', binwidth = 0.5) +
  coord_cartesian(xlim = c(-20, 45)) +
  ggtitle("Mean = 28 and Standard deviation = 1")


grid.arrange(p5, p6)
```

#### Watch the scales

```{r}
norms_4 <- tibble(x_wider = rnorm(1000, mean = 15, sd = 10), 
                  x_narrower = rnorm(1000, mean = 28, sd = 1))


p7 <- ggplot(norms_4, aes(x = x_wider)) +
  geom_histogram(fill = 'skyblue2', colour = 'black', binwidth = 5) +
  ggtitle("Mean = 15 and Standard deviation = 10")


p8 <- ggplot(norms_4, aes(x = x_narrower)) +
  geom_histogram(fill = 'skyblue2', colour = 'black', binwidth = 0.5) +
  ggtitle("Mean = 28 and Standard deviation = 1")


grid.arrange(p7, p8)
```

## Calculating our statistics

### Load and clean some data

```{r}
wages <- read_csv("http://vincentarelbundock.github.io/Rdatasets/csv/Ecdat/Bwages.csv") %>% 
  select(-X1, -sex) %>% 
  mutate(educ = factor(educ)) 

wages
```

2. Using the formula noted earlier, calculate the variance using 'dplyr' verbs for the 'wage' variable.

```{r}
wages %>% select(wage) %>% 
  mutate(avg_wage = mean(wage), 
         wage_minus_avg_wage = wage - avg_wage, 
         sqr_wage_minus_avg_wage = wage_minus_avg_wage**2) %>%
  summarise(var_wage = sum(sqr_wage_minus_avg_wage) / (n() - 1))
```

3. Using R functions and 'summarize', calculate the mean, standard deviation, median, and IQR directly for the 'wage' variable.

```{r}
wages %>% summarise(mean_wage = mean(wage), sd_wage = sd(wage), median_wage = median(wage), IQR_wage = IQR(wage))
```

## Why visualize when we have statistics?

Let’s take a look at a famous group of 4 data sets called Anscombe’s quartet.

```{r}
anscombe
```

4. Calculate and print the mean of 'x1', 'x2', 'x3', and 'x4'.

```{r}
anscombe %>% select(contains('x')) %>% 
  summarise_all(funs(mean))

```

5. Calculate and print the standard deviation of 'x1', 'x2', 'x3', and 'x4'.

```{r}
anscombe %>% select(contains('x')) %>% 
  summarise_all(funs(sd))
```

6. Calculate and print the mean of 'y1', 'y2', 'y3', and 'y4'.

```{r}
anscombe %>% select(contains('y')) %>% 
  summarise_all(funs(mean))
```

7. Calculate and print the standard deviation of 'y1', 'y2', 'y3', and 'y4'.

```{r}
anscombe %>% select(contains('y')) %>% 
  summarise_all(funs(sd))
```

Now, let’s put the data into tidy format and calculate all at once.

8. Use 'anscombe_tidy' to calculate the mean and standard deviation of 'x' and 'y' per 'set'.

```{r}
anscombe_x <- anscombe %>% select(x1, x2, x3, x4) %>% 
                gather(set_x, x, x1:x4) %>% 
                mutate(set_x = str_replace(set_x, 'x', ''))

anscombe_y <- anscombe %>% select(y1, y2, y3, y4) %>% 
                gather(set_y, y, y1:y4) %>% 
                mutate(set_y = str_replace(set_y, 'y', ''))

anscombe_tidy <- cbind(anscombe_x, anscombe_y) %>%
                    select(-set_x) %>% 
                    rename(set = set_y) 

anscombe_tidy %>% 
  group_by(set) %>%
  summarise(mean_x = mean(x), sd_x = sd(x), mean_y = mean(y), sd_y = sd(y))
```

9. Using the 'anscombe_tidy' create a scatter plot of x versus y but facet on the 'set' variable.

```{r}
ggplot(anscombe_tidy, aes(x = x, y = y)) + geom_point() + facet_wrap(~set)
```

10. Does the above plot give you an idea as to why we would still visualize data even though we can compute summary statistics?

Yes. The plots show that similar summary statistics can arise from data with very different distributions.
