---
title: "Introduction to linear regression - Part 1"
output:
  html_document:
    highlight: pygments
    theme: cerulean
  pdf_document: default
Name: THI NGOC DIEM NGUYEN
ID: 0778061
Week: 10
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

## Batter up 

The movie [Moneyball](http://en.wikipedia.org/wiki/Moneyball_(film)) focuses on
the "quest for the secret of success in baseball". It follows a low-budget team, 
the Oakland Athletics, who believed that underused statistics, such as a player's 
ability to get on base, betterpredict the ability to score runs than typical 
statistics like home runs, RBIs (runs batted in), and batting average. Obtaining 
players who excelled in these underused statistics turned out to be much more 
affordable for the team.

In this lab we'll be looking at data from all 30 Major League Baseball teams and
examining the linear relationship between runs scored in a season and a number 
of other player statistics. Our aim will be to summarize these relationships 
both graphically and numerically in order to find which variable, if any, helps 
us best predict a team's runs scored in a season.

## Load packages

```{r packages, warning = FALSE, message = FALSE}
library(tidyverse)
```

## The data

Let's load up the data for the 2011 season.

```{r load-data}
download.file("http://www.openintro.org/stat/data/mlb11.RData", destfile = "mlb11.RData")
load("mlb11.RData")
mlb11
```

In addition to runs scored, there are seven traditionally used variables in the 
data set: at-bats, hits, home runs, batting average, strikeouts, stolen bases, 
and wins. There are also three newer variables: on-base percentage, slugging 
percentage, and on-base plus slugging. For the first portion of the analysis 
we'll consider the seven traditional variables. At the end of the lab, you'll 
work with the newer variables on your own.

1.  What type of plot would you use to display the relationship between `runs` 
    and one of the other numerical variables? Plot this relationship using the 
    variable `at_bats` as the predictor. Does the relationship look linear? If 
    you knew a team's `at_bats`, would you be comfortable using a linear model 
    to predict the number of runs?

> I would use a scatter plot to display the relationship between `runs` and another numerical variables.

> Since `at_bats` is the predictor variable (the explanatory variable), then the scatter plot should have `at_bats` on the x axis and `runs` (the response variable) on the y axis.

```{r}
ggplot(mlb11, aes(x = at_bats, y = runs)) + 
  geom_point()
```

> The relationship looks linear.

> Yes, I would you be comfortable using a linear model to predict the number of runs. It's not a perfect relationship but it seems like it would do a pretty good job predicting `runs`.

If the relationship looks linear, we can quantify the strength of the
relationship with the correlation coefficient.

```{r cor}
cor(mlb11$runs, mlb11$at_bats)
```

## Sum of squared residuals

Think back to the way that we described the distribution of a single variable. 
Recall that we discussed characteristics such as center, spread, and shape. It's
also useful to be able to describe the relationship of two numerical variables, 
such as `runs` and `at_bats` above.

2.  Looking at your plot from the previous exercise, describe the relationship 
    between these two variables. Make sure to discuss the form, direction, and 
    strength of the relationship as well as any unusual observations.

Just as we used the mean and standard deviation to summarize a single variable, 
we can summarize the relationship between these two variables by finding the 
line that best follows their association. 

You will use the following function to select the line that you think does 
the best job of going through the cloud of points.

```{r}
# This function takes in a value for the slope and intercept of a line and plots
# the original data points plus the specified line. It also shows the residuals 
# and calculates the sum of the squares of the residuals and prints that in the title.

# We can think of any line as a model for the relationship between runs and at_bats,
# although some of these models (lines) will be much worse representations of our 
# data than others. This function allows you to explore finding the line that 
# minimizes the sum of the squared residuals throught trial and error. 

best_line <- function(slope, intercept, residuals = TRUE) {
  mlb <- mlb11 %>% mutate(y_predicted = intercept + slope * at_bats,  
                          square_residuals = (runs - y_predicted)^2)

  ssr <- mlb %>% summarize(ssr = round(sum(square_residuals), 0))
  
  if (residuals) {
    p <- ggplot(mlb, aes(x = at_bats, y = runs)) +
          geom_point(shape = 21, fill = 'skyblue', size = 2) + 
          geom_line(aes(x = at_bats, y = y_predicted), colour = 'blue') + 
          geom_segment(aes(x = at_bats, y = runs, 
                           xend = at_bats, yend = y_predicted), linetype = 2) +
          ggtitle(paste0("Sum of Squared Resdiduals: ", ssr[[1]]))
  } else {    
    p <- ggplot(mlb, aes(x = at_bats, y = runs)) +
          geom_point(shape = 21, fill = 'skyblue', size = 2) + 
          geom_line(aes(x = at_bats, y = y_predicted), colour = 'blue')
  }
  p
}
```

You can use the `best_line` function above to plot the line you want 
by specifying the slope and intercept. The line you specify will be shown in blue and 
the residuals will be noted by dotted black lines. Note that there are 30 residuals, 
one for each of the 30 observations. Recall that the residuals are the difference between 
the observed values and the values predicted by the line:

\[
  e_i = y_i - \hat{y}_i
\]

The most common way to do linear regression is to select the line that minimizes
the sum of squared residuals. To visualize the squared residuals, their sum has 
been added as the title of plot.

3.  Using `best_line`, choose a line that does a good job of minimizing the sum of
    squares as well as doing the best job of going through the cloud of points. 
    Run the function several times. What was the smallest sum of squares that you got? 
    How does it compare to your neighbors? Record your best values for the *slope*, 
    *intercept*, and sum of squared residuals. 

```{r}
best_line(slope = 0.545, intercept = -2300)
```

>I ran the `best_line()` function several times and decided that the best fit line has a slope of 0.545, an intercept of -2300, and a sum of squared residuals of 133450.

> I noticed that the sum of squared residuals went down the more my line went through the 'middle' of the data points.

## The linear model

It is rather cumbersome to try to get the correct least squares line, i.e. the 
line that minimizes the sum of squared residuals, through trial and error. 
Instead we can use the `lm` function in R to fit the linear model (also known as a 
regression line).

```{r m1}
m1 <- lm(runs ~ at_bats, data = mlb11)
```

The first argument in the function `lm` is a formula that takes the form 
`y ~ x`. Here it can be read that we want to make a linear model of `runs` as a 
function of `at_bats`. Or, that `at_bats` is the *explanatory* variable and `runs` 
is the *response* variable. The second argument specifies that R should look in the 
`mlb11` data frame to find the `runs` and `at_bats` variables.

The output of `lm` is an object that contains all of the information we need 
about the linear model that was just fit. We can access this information using 
the summary function.

```{r summary-m1}
summary(m1)
```

Let's consider this output piece by piece. First, the formula used to describe 
the model is shown at the top. After the formula you find the five-number 
summary of the residuals. The "Coefficients" table shown next is key; its first 
column displays the linear model's y-intercept and the coefficient of `at_bats`.
With this table, we can write down the least squares regression line for the 
linear model:

\[
  \hat{y} = -2789.2429 + 0.6305 * at\_bats
\]

One last piece of information we will discuss from the summary output is the 
Multiple R-squared, or more simply, $R^2$. The $R^2$ value represents the 
proportion of variability in the response variable that is explained by the 
explanatory variable. For this model, 37.3% of the variability in `runs` is 
explained by `at-bats`.

4.  How close were your estimates for the *slope* and *intercept* of your best fit 
    line to the values produced by the `lm` function? 
    
> The slope of my best fit line (from trial and error using the `best_line()` function, is 0.545 while the slope for the linear model (from least squares algorithm using the 'lm()' function) is 0.631. My value is about 86% of the model value.

> The intercept of my best fit line is -2300 while the intercept for the linear model is -2789. This is about 82% of the model value.
    
5.  Fit a new model that uses `homeruns` to predict `runs`. Using the estimates 
    from the R output, write the equation of the regression line. What does the 
    slope tell us in the context of the relationship between success of a team 
    and its home runs?

```{r}
# model with `homeruns` as the explanatory variable
m2 <- lm(runs ~ homeruns, data = mlb11)

# output from the model
summary(m2)
```

> The equation for the regression line is:

\[
  \hat{runs} = 415.2389 + 1.8345 * homeruns
\]

> A slope 1.8345 means that for every 'homerun', on average, a team will have 1.835 more 'run'. This would imply that a team with more homeruns is more likely to be successful since they would be more likely to have more runs.

## Prediction and prediction errors

6. Use output from `lm` and the `best_line` function to create a scatter plot 
with the least squares line plotted with the data points. (Be sure to set `residuals = FALSE` 
when you call the function this time.) 

```{r reg-with-line}
best_line(slope = 0.6305, intercept = -2789.2429, residuals = FALSE)
```

> The main goal here is to visualize your linear model along with the actual data. This could be accomplished using 'geom_abline()':

```{r}
ggplot(mlb11, aes(x = at_bats, y = runs)) + 
  geom_point() +
  geom_abline(slope = 0.6305, intercept = -2789.2429)
```

> Note that this is also the same as the output from 'geom_smooth()':

```{r}
ggplot(mlb11, aes(x = at_bats, y = runs)) + 
  geom_point() +
  geom_abline(slope = 0.6305, intercept = -2789.2429) +
  geom_smooth(method = 'lm', se = FALSE, colour = 'orange')
```

This line can be used to predict $y$ for any value of $x$. When 
predictions are made for values of $x$ that are beyond the range of the observed
data, it is referred to as *extrapolation* and is not usually recommended. 
However, predictions made within the range of the data are more reliable. 
They're also used to compute the residuals.

7.  If a team manager saw the least squares regression line and not the actual 
    data, how many runs would he or she predict for a team with 5,578 at-bats? 
    Is this an overestimate or an underestimate, and by how much? In other 
    words, what is the residual for this prediction?

> The team manager would predict '-2789.2429 + 0.6305 * 5578' runs.

> To calculate the residual we use the formula: $e_i = y_i - \hat{y}_i$, where $y_i$ is the actual value from the data and $\hat{y}_i$ is the corresponding value predicted by our linear model.

> To determine if the prediction that the team manager made above is an overestimate or an underestimate, we need to calculate the residual for this prediction. Let's see if we have an actual value for 'runs' when 'at_bats = 5578'.

```{r}
mlb11 %>% filter(between(at_bats, 5575, 5580))
```

> The closet value in our actual data corresponds to 'at_bats = 5579' so we will use this value to calculate the residual.

> $\hat{y}_i = 727.6861$, corresponding to the value our linear model predicts when 'at_bats = 5578'.

> $y_i = 713$, corresponding to the value for 'runs' in our acutal data when 'at_bats = 5579' (the closet value to 5578).

> Thus, $e_i = 713 - 727.6861 = -14.6861$

> Since the residual for this point is negative, that means the value predicted by our linear model is an 'overestimate'.

<div id="license">
This a modified version of a product of OpenIntro that is released under a [Creative Commons 
Attribution-ShareAlike 3.0 Unported](http://creativecommons.org/licenses/by-sa/3.0). 
This lab was adapted for OpenIntro by Andrew Bray and Mine &Ccedil;etinkaya-Rundel 
from a lab written by the faculty and TAs of UCLA Statistics.
</div>